---
title: "Project 3"
format: html
editor: visual
---

## Project 3 Code Base Submission

------------------------------------------------------------------------

### Group Members:

Long Lin, Emily Elmouaquite, ZIHAO YU, Pascal Hermann Kouogang Tafo

------------------------------------------------------------------------

**First Data Source**: https://www.kaggle.com/datasets/joerakhimov/data-scientist-skills

https://raw.githubusercontent.com/longflin/DATA-607-Project-3/refs/heads/main/data_scientist_skills.csv

**Second Data Source**: https://www.kaggle.com/datasets/fahadrehman07/data-science-jobs-and-salary-glassdoor?select=glassdoor_jobs.csv

https://raw.githubusercontent.com/longflin/DATA-607-Project-3/refs/heads/main/Glassdoor_Salary_Cleaned_Version.csv

------------------------------------------------------------------------

## First Data Source

For the first data source, we grabbed a dataset from kaggle on data scientist skills.

We created a data frame from first data source by reading in the .csv file.

```{r}
library(tidyverse)

url <- "https://raw.githubusercontent.com/longflin/DATA-607-Project-3/refs/heads/main/data_scientist_skills.csv"

ds_skills <- read_csv2(
  file = url,
  show_col_types = FALSE,
  progress = FALSE
)

head(ds_skills)
```

Next, we transformed the data from a wide data format to a long data format using pivot_longer on the columns pertaining to skills.

```{r}
ds_skills_long <- ds_skills |>
  pivot_longer(
    cols = c(Statistics:Azure),
    names_to = "Skills",
    values_to = "Value"
  )

head(ds_skills_long)
```

Next, we removed the columns with information that we were not interested in using the select function.

```{r}
library(dplyr)
# Remove unncessary columns from ds_skills_long
ds_subset <- select(ds_skills_long, -Platform, -Bachelor, -Masters, -PhD, -Communication, -English, -German)
head(ds_subset)
```

Next, we removed the rows that had a `0` in the `Value` column because the job did not require the skill.

```{r}
ds_subset_remove_0s <- ds_subset[ds_subset$Value != 0, ]
head(ds_subset_remove_0s)
```

Next, we removed the `Value` Column altogether because all of the values in that column were `1`.

```{r}
ds_subset_remove_value_column <- select(ds_subset_remove_0s, -Value)
head(ds_subset_remove_value_column)
```

Next, we added a column named `Job Title` and assigned a value of `Data Scientist` to each of the rows because the data source is about data scientist jobs.

```{r}
ds_job_title <- ds_subset_remove_value_column |>
  mutate(
    `Job Title` = "Data Scientist"
  ) %>%
  relocate(`Job Title`, .before = 1)

head(ds_job_title)
```

## Second Data Source

For the second data source, we created a dataframe from the .csv file.

```{r}
library(tidyverse)

url2 <- "https://raw.githubusercontent.com/longflin/DATA-607-Project-3/refs/heads/main/Glassdoor_Salary_Cleaned_Version.csv"

glassdoor_skills <- read_csv(
  file = url2,
  show_col_types = FALSE,
  progress = FALSE
)

head(glassdoor_skills)
```

Next, we renamed the columns in the data frame to match with the first data frame's format for the same skills.

```{r}
glassdoor_skills_rename_columns <- rename(glassdoor_skills, Python = python_yn, R = R_yn, Spark = spark, AWS = aws, Excel = excel)

head(glassdoor_skills_rename_columns)
```

Next, we used the `pivot_longer` function to convert the data frame from a wide format to a long format.

```{r}
glassdoor_skills_long <- glassdoor_skills_rename_columns |>
  pivot_longer(
    cols = c(Python:Excel),
    names_to = "Skills",
    values_to = "Value"
  )

head(glassdoor_skills_long)
```

Next, we removed the columns that we were not interested in using the `select` function.

```{r}
library(dplyr)
# Remove unncessary columns
glassdoor_subset <- select(glassdoor_skills_long, -"Salary Estimate", -"Job Description", -Rating, -Headquarters, -Size, -Founded, -"Type of ownership", -Industry, -Sector, -Revenue, -Competitors, -hourly, -employer_provided, -min_salary, -max_salary, -avg_salary, -company_txt, -job_state, -same_state, -age)

head(glassdoor_subset)
```

Next, we removed the rows that had a `Value` of `0`.

```{r}
glassdoor_subset_remove_0s <- glassdoor_subset[glassdoor_subset$Value != 0, ]

head(glassdoor_subset_remove_0s)
```

Next, we removed the `Value` column because the only rows remaining had a `Value` of `1`.

```{r}
glassdoor_remove_value_column <- select(glassdoor_subset_remove_0s, -Value)

head(glassdoor_remove_value_column)
```

Next, we filtered jobs based on if they had the word `data` in the `Job Title` or not, in order to get relevant job listings. We did this using the `stringr` library and the `str_detect` function.

```{r}
library(stringr)
glassdoor_remove_value_column$`has_data_in_name` <- str_detect(glassdoor_remove_value_column$`Job Title`, "Data")

head(glassdoor_remove_value_column)
```

Next, we removed the columns that did not have the word `data` in the `Job Title`.

```{r}
glassdoor_data_jobs_remove_false <- glassdoor_remove_value_column[glassdoor_remove_value_column$has_data_in_name != FALSE, ]

head(glassdoor_data_jobs_remove_false)
```

Next, we removed the `has_data_in_name` column because the value is `TRUE` for all of the rows since we removed the columns that had a value of `FALSE`.

```{r}
glassdoor_final <- select(glassdoor_data_jobs_remove_false, -has_data_in_name)

head(glassdoor_final)
```

## Join Data Frames

Before joining the data frames, we made the format of the first data frame similar to that of the second data frame.

```{r}
ds_final <- rename(ds_job_title, "Company Name" = Company, Location = Country)
head(ds_final)
```

Next, we joined the two data frames using the `dplyr` library and the `rbind` function.

```{r}
library(dplyr)

combined_df <- rbind(ds_final, glassdoor_final)

head(combined_df)
colnames(combined_df)

```

## Database

```{r}
#Create database tables
#Job
jobTable <- combined_df %>%
  distinct(`Job Title`, `Company Name`, Location) %>%
  mutate(jobId = row_number()) %>%
  rename(jobName = `Job Title`)
#Skill
skillTable <- combined_df %>%
  distinct(Skills) %>%
  mutate(skillId = row_number()) %>%
  rename(skillName = Skills)
#jobSkill
jobSkillTable <- combined_df %>%
  left_join(jobTable,
            by = c("Job Title" = "jobName",
                   "Company Name",
                   "Location")) %>%
  left_join(skillTable,
            by = c("Skills" = "skillName")) %>%
  select(jobId, skillId) %>%
  distinct()

#Load database/ SQL libraries
library(DBI)
library(RSQLite)

#Connect to database
con <- dbConnect(SQLite(), "ds_jobs.db")

#Create db tables
dbExecute(con, "
CREATE TABLE job (
  jobId INTEGER PRIMARY KEY,
  jobName TEXT,
  location TEXT
);
")
dbExecute(con, "
CREATE TABLE skill (
  skillId INTEGER PRIMARY KEY,
  skillName TEXT
);
")
dbExecute(con, "
CREATE TABLE jobSkill (
  jobId INTEGER,
  skillId INTEGER,
  PRIMARY KEY (jobId, skillId),
  FOREIGN KEY (jobId) REFERENCES Job(jobId),
  FOREIGN KEY (skillId) REFERENCES Skill(skillId)
);
")

#Populate tables with data
dbWriteTable(con, "job", jobTable %>% select(jobId, jobName, Location), append = TRUE)
dbWriteTable(con, "skill", skillTable %>% select(skillId, skillName), append = TRUE)
dbWriteTable(con, "jobSkill", jobSkillTable, append = TRUE)
```

```{r}
#Check
DBI::dbListTables(con)
DBI::dbReadTable(con, "job")
DBI::dbReadTable(con, "skill") 
DBI::dbReadTable(con, "jobSkill") 
```

## Analysis

```{r}
combined_df %>%
  count(Skills, sort = TRUE) %>%
  head(10) %>%
  ggplot(aes(x = reorder(Skills, n), y = n, fill = Skills)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = n), 
            hjust = -0.05,
            size = 4,
            color = "black") +
  coord_flip() +
  labs(title = "Top 10 Most Demanded Skills", x = "Skill", y = "Frequency")+
  theme_minimal()
```

Based on the plot above for the combined data set, we're able to see the top 10 skills that are in demand. Python is the clear leader with Excel and Spark as the second and third most in demand, respectively.
